---
title: "Reproducible Research: Peer Assessment 1"
author: "Robert Monné"
date: "7 december 2014"
output: 
  html_document:
    keep_md: true
---

This is a markdown document to provide answers to the Peer Assignment for the course Reproducible research in the data science specialisation

## Loading and preprocessing the data
We will first load and preprocess the data from the .csv file. We use the lubridate package to transform the dates to a more usable format

```{r load_data}
data <- read.csv("activity.csv")
library(lubridate)
data$date <- ymd(data$date)
```

## What is mean total number of steps taken per day?
Make a histogram of the total number of steps taken each day  
And compute the mean and median of the total number of steps taken each day  

```{r explore_data}
totalsteps <- tapply(data$steps, data$date, sum)
hist(totalsteps, xlab="Total steps per day", main="Histogram of Total steps per day")
mean(totalsteps)
median(totalsteps)

```


## What is the average daily activity pattern?
Calculate the average of steps per time interval  
Plot the intervals on the X-axis and the Average amount of steps on the Y-axis  
In what interval lies the maximum of average steps?  

```{r average}
averagesteps <- as.data.frame(tapply(data$steps, data$interval, mean, na.rm=T))

plot(rownames(averagesteps), averagesteps[,1], type = "l", xlab="Interval", ylab="Average Steps")

names(which.max(averagesteps[,1]))

```




## Imputing missing values
First calculate how many missing values are in the current dataset

Then we use a for loop to set the NAs with a new value (the mean of the steps on the given interval)


```{r missing values, cache=TRUE}
sum(is.na(data))
newdata <- data

#for every row in the dataset
for(i in 1:length(newdata[,1])) {
  #check if there is a NA
  if(is.na(newdata[i,1])) {
    #grab the associated interval of the NA
    interval <- newdata[i,3]
    #use the found interval to grab the mean steps of that interval and set is as new value for the NA
    #we grab the first value resulting vector due to overlappings like: 815 & 1815
    newdata[i,1] <- averagesteps[grep(as.character(interval), rownames(averagesteps)),1][1]
  }
}

sum(is.na(newdata))

totalstepsnew <- tapply(newdata$steps, newdata$date, sum)
hist(totalstepsnew, xlab="Total steps per day", main="Histogram of Total steps per day")
mean(totalstepsnew)
median(totalstepsnew)

```

## Are there differences in activity patterns between weekdays and weekends?




